Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-10 intro.tex
+
lots of comments

At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations (addresses) seen by most code are not, in fact, the exact location in physical memory where data reside. Instead these are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS. This is what enables separation of process memory resources:
the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted to it by the kernel will fail.
But this is complicated by the fact that the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; and the fact that control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, which the kernel still accesses via virtual address, leading to indirect cycles.

Further complicating matters is that addresses themselves bear no information about which address space they originate from. For user processes this is of little concern, as these have access to only their own address space. But the kernel has access to all address spaces. Mixing up addresses from different address spaces leads to severe bugs. More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in multiple but still not all.

This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor. 
We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid modal logic, which can label assertions true under \emph{other} circumstances (i.e., in another address space) with a modality indexed by a name for that space (in our case, the root of the page tables for an address space). This offers a \textit{convenient} and \textit{powerful way to modularly}
\begin{itemize}
\item isolate assertions specific to a particular address space,
\item explicitly state when an assertion is true across address spaces,
\item manipulate address spaces from within other address spaces, and
\item reason about change in address spaces.
\end{itemize}
  These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, which was only able to work with a single address space (the current address space on the CPU).

\begin{figure}
   \includegraphics[width=0.5\columnwidth]{architecture.png}
  \caption{Layers of Implementation}
  \label{fig:architecture}
\end{figure}
We realize and show our reasoning principles, showin in Figure \ref{fig:architecture}
\begin{itemize}
\item we develop these ideas in an Iris-based\mytodo{cite JFP} separation logic, a logic for address-virtualization (\textsf{vProp} in Figure \ref{fig:architecture})
\item we model \textsf{AMD64} in Coq, and specify and prove its instructions in \textsf{vProp}
\item and verify simplified versions of several critical virtual-memory-related pieces of OS functionality, including mapping and unmapping pages, and describing the self-referential identity maps used in most kernels to allow convenient access to physical memory.
\end{itemize}

>>>>

At its core, the virtual memory functionality of modern CPUs is about \emph{location virtualization}: the memory locations
(addresses) seen by most code are not, in fact, the exact location in physical memory where data reside. Instead these 
are \emph{virtual} addresses, which are mapped to actual physical resources by the cooperation of the hardware and OS. 
This is what enables separation of process memory resources:
the OS manipulates hardware functionality to ensure that any attempt by a process to access memory not explicitly granted 
to it by the kernel will fail. But this is complicated by the fact that 
%the OS and hardware can also enable, shared (overlapping) access to physical memory regions; the fact that the kernel data structures themselves are accessed via virtual memory addresses; 
%and the fact that 
control over these mappings of virtual to physical addresses is itself mediated by \emph{in-memory data structures}, 
which the kernel still accesses via virtual address, leading to indirect cycles.

Further complicating matters is that addresses themselves bear no information about which address space they originate 
from. For user processes this is of little concern, as these have access to only their own address space. But the kernel has
(or can grant itself) access to all address spaces. Mixing up addresses from different address spaces leads to severe bugs.
More concerning, keeping track of which \emph{assertions} hold in different address spaces during kernel verification is 
difficult: some assertions should hold across all address spaces, while others hold in only one, and others may hold in 
multiple but still not all.

This kind of context-dependent assertion, where a fact may be true in one address space but not others, has a modal flavor. 
We propose tackling the verification of virtual memory subsystems (and kernels more broadly) by adapting ideas from hybrid
modal logic, which can label assertions true under \emph{other, named} circumstances (i.e., in another address space) with a 
modality indexed by a name for that space (in our case, the root of the page tables for an address space). This offers a 
\textit{convenient} and \textit{powerful} way to \emph{modularly}
\begin{itemize}
\item isolate assertions specific to a particular address space,
\item explicitly state when an assertion is true across address spaces,
\item manipulate address spaces from within other address spaces, and
\item reason about change in address spaces.
\end{itemize}
These advantages make this approach to reasoning about virtual memory more flexible than prior program logic techniques~\cite{kolanski08vstte,kolanski09tphols}, 
which were only able to work with a single address space (the current address space on the CPU) because they were unable
to speak directly \emph{within the logic} about other address spaces.

\begin{figure}
   \includegraphics[width=0.5\columnwidth]{architecture.png}
  \caption{Layers of Implementation}
  \todo[inline]{Colin says: I don't think this figure helps readers; it is too abstract}
  \label{fig:architecture}
\end{figure}

We realize and show our reasoning principles, shown in Figure \ref{fig:architecture}
\begin{itemize}
\item We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
      implemented as an embedded separation logic within Iris~\cite{jung2018iris} separation logic.
\item We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
\item We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
      including mapping  pages, the assembly code for switching address spaces, and code which manipulates page
      tables for \emph{currently mapped memory} without disturbing mappings.\todo[inline]{Colin says: I have optimistically
      added this last example to the text here, because I think we can probably do it. Leave this TODO
      here until we either do or cut it.}
      The latter in particular goes beyond the capabilities of prior work.
\end{itemize}

### 2023-04-10 programlogic.tex
small stuff on figures
some cr3 removal from defs
+
lots of comments

Our reasoning principles are shaped around understanding the following perspectives on the vritual-memory abstraction:

@@

\item \textit{sharing}: a bag of physical pages are reached through a set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses. This sharing imposes constraints on giving a definition to the virtual-address in terms of physical (L4-L1) page-table memory accesses
\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} on its address-space. 
\item \textit{address-spaces as modal-contexts}: we call these context-agnostic resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}

@@

Based on these perspectives, we coneptualize truth on an address-space through the contingency it exhibits: \textit{assertion P holds on this address-space indexed by its roots address}. Therefore, as a first step, we lift standard bi proposition to be indexed with $\mathcal{W}_{64}$ as shown in Figure \ref{fig:vprop}, and, most notably, all abstractions we explain in this section as part of our custom-tailored address-space logic are type $\textsf{vProp }\Sigma$.
\begin{figure}[!ht]

@@

\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{def}{=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma

@@

\subsection{Points-To Assertions}
\label{sec:pointsto}
We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
\item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
\item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
\paragraph{Register points-to} The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$. The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.
\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g. \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.  
\begin{figure}[!ht]
\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{def}{=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
\caption{Physical Points-to with Nested Masking}
  \label{fig:physicalpointsto}
Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L4translate}})} \quad
      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr
  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
    \[\begin{array}{l}
    \hbox{(\TirNameStyle{{L3translate}})} \quad
    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr
  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
    \[ \begin{array}{l}
            \hbox{(\TirNameStyle{{L2translate}})} \quad 
       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr
  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L1translate}})} \quad
      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr
  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr

\end{itemize}
\begin{remark}[A Strong Definition for Virtual Memory Addressing]
  \label{rem:strongvmem}
  One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation (Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, we can obtain and update the value mapped value at this physical page address through our virtual memory address.

\begin{figure*}
  \[
  \begin{array}{l}
\begin{array}{l}
  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{def}{:=}  \lambda \crthree. \;\; \\
  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp

  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
 \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
\end{array} \\
\\
\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
 \qquad\qquad \nfpointsto{\mask\vaddr\ft\crthree}{\mask\vaddr\tw\crthree}\entryf\qfracfotsss\naddr \ast \\ 
 \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
   \end{array}
\]
\caption{A Strong Virtual Points-to Relation}
\todo[inline]{These fractions aren't quite right (though I see you added the variance between levels), I can walk you through in our meeting tomorrow.}
  \label{fig:strongvirtualpointsto}
\end{figure*}
\end{remark}
%then we reach to a physical page addr (e.g. \textsf{pa}) 
%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]

  \subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} is too strong to specify virtual memory operations updating a page table entry. To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) \textit{only} assure the existence of the certain number of mappings -- 512 at each level -- abstracting virtual address translations, and updating any of these mappings would make the other mappings, address translations sharing the page table mappings. Therefore, we need to consturct logical abstractions ensuring sound updates to the page-table entries in the existence of sharing.
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
    \exists \entryo \ldotp
    \sumapaces\rtv\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}  
  One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation. To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$. We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
 where
  \begin{itemize}
  \item we decorate the fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$) with the physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
  \item we place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership grants sound update on the page table entries
  \end{itemize}

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
\end{array}
\]
\caption{Global Address-Space Invariant}
  \label{fig:peraspaceinvariant}
  \end{figure*}

@@

\[
  \begin{array}{l}
    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
         [r](P) \; \mathsf{vProp} \; \Sigma \; := \; \lambda \mathsf{P},\; \mathsf{P r}. 
  \end{array}
  \]
  \label{fig:modaldef}
  \end{figure}

>>>>

Our reasoning principles derive from the following perspectives on the virtual-memory abstraction:

@@

\item \textit{sharing}: the physical location backing a virtual address is located (during a page table walk) through a 
      set of physical page-table (L4-L1 page-tables) acceses that are shared amongst different virtual addresses (specifically,
      those on the same page of memory). This sharing imposes constraints on giving a definition to the virtual-address 
      in terms of physical (L4-L1) page-table memory accesses
\item \textit{context-agnostic-resources}: each virtual address is valid under a certain address-space, but it does not represent this \textit{knowledge} of its address-space. 
\item \textit{address-spaces as modal contexts}: we call these context-dependent\todo{Colin switched away from `agnostic', which is the opposite!} resources (virtual-address pointstos) as \textit{context-resource}, and the context identfied by a unique \textit{namespace}, and  defining the truth for these context resources as \text{modal-context}

@@

Based on these perspectives, we coneptualize truth in an address-space through the contingency it exhibits: 
\textit{assertion P holds only relative to a certain address space}. 
Therefore, as a first step, we work in a pointwise lifting of Iris's standard BI proposition,
essentially working with separation logic assertions indexed by a page table root as a $\mathcal{W}_{64}$
(Figure \ref{fig:vprop}), which we call $\textsf{vProp }\Sigma$; this is the (\textsc{Coq}) type of assertions in our logic.
\begin{figure}[t]

@@

\textsf{vProp } \Sigma \; : \; \textsf{bi} \; \stackrel{{\triangle}}{:=} \mathcal{W}_{64} \rightarrow_{\textsf{b}} \textsf{iPropI } \Sigma

@@

\subsection{Points-To Assertions}
\label{sec:pointsto}
\todo[inline]{Colin says: Current limit of \emph{thorough} revisions}
\todo[inline,color=red]{Colin says: This section at a glance looks on the right track, but none of the assertions are actually
written as vProps, they're all still iProps. I can fix that as I go}
\todo[inline]{Colin says: There's really no need to discuss the 52-12 split in the physical pointsto definition.
That level of detail of the operational semantics \emph{is} too detailed for the paper (but should be mentioned in the
artifact docs). I also don't know what the `ent' and `rtv' and similar are in those split physical pointstos}
We abstract physical memory addressing and registers naming with well-known separation logic assertion \textit{points-to} assertions.
\begin{enumerate}
\item Physical address points-to, $\pfpointsto\locsf\vpts\qfrac\ppts$
\item Register points-to, $\pfpointsto\rg\rv\qfrac\rpts$
\end{enumerate}
\paragraph{Register points-to} The assertion $\pfpointsto\rg\rv\rpts\qfrac$ ensures the ownership of the register $\rg$ naming the register value $\rv$. The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping, and grants update permission on it, otherwise, any value $0 < \qfrac <1$ represents partial ownership granting readonly permission on the mapping.

\todo[inline]{We should drop all talk about PDE, PDPT, etc., and just use L1-L4}
\paragraph{Physical-Memory points-to} Our physical memory points-to relation in Figure \ref{fig:physicalpointsto} for a page table entry (e.g. \textsf{PDE} in Figure \ref{fig:pagetables}) abstracts masking -- $|^{52}$ for frame for and $|^{12}$ for offset -- through logical nested maps.  
\begin{figure}[t]
\[
\begin{array}{cl}
\pfpointsto\locsf\vpts\qfrac\ppts \stackrel{{\triangle}}{:=} & \nfpointsto{\mask\locsf\ft\generalentry}{\mask\locsf\tw\generalentry}\vpts\qfrac\naddr
\end{array}
\]
\caption{Physical Points-to with Nested Masking}
  \label{fig:physicalpointsto}
\end{figure}
Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:
\begin{itemize}
  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks (\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L4translate}})} \quad
      \exists \entryf \ldotp \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr
      \end{array}
      \]
  \item Page Directory Pointer Level Translation (L3): Performs the second level address translation to get the PDP entry (PDPE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PLM4E}) to get the starting address of the PDP table (PDPT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PLM4E})
    \[\begin{array}{l}
    \hbox{(\TirNameStyle{{L3translate}})} \quad
    \exists \entrytr \ldotp \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr
    \end{array}
    \]
  \item Page Directory Table Level Translation (L2): Performs the second level address translation to get the PD entry (PDE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDPE}) to get the starting address of the PD table and using the offset mask \textsf{va}$|^{12}$(\textsf{PDPE})
    \[ \begin{array}{l}
            \hbox{(\TirNameStyle{{L2translate}})} \quad 
       \exists \entrytw \ldotp \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr
      \end{array}
      \]
  \item Page Table Level Translation (L1): Performs the second level address translation to get the PT entry (PTE) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PDE}) to get the starting address of the Page table (PT) and using the offset mask \textsf{va}$|^{12}$(\textsf{PDE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{L1translate}})} \quad
      \exists \entryo \ldotp \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr
      \end{array}
      \]
  \item Page Address Level Translation: Performs the second level address translation to get the physical page address in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{PTE}) to get the starting address of the physical page address table and using the offset mask \textsf{va}$|^{12}$(\textsf{PTE})
    \[ \begin{array}{l}
      \hbox{(\TirNameStyle{{PageLebelAccess}})} \quad
      \exists \vpage \ldotp \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
      \end{array}
      \]

\end{itemize}
\subsection{An Overly-Restrictive Definition for Virtual Memory Addressing}
One could give a naive logical definition to the virtual mememory addressing as conjuctions of each level translation 
(Figure \fref{fig:strongvirtualpointsto}). The very first assurance our logical constructions need to be able assert is 
the \textit{existence} of the page table walk. Our naive attempt shown in Figure \ref{figure:strongvirtualpointsto}, 
we see that a virtual address with a certain fractional permissions at each level of the page-table walk gives us sound 
assurance of the path existence reaching to a page table entry, consequently, with a full-ownership of page address, 
we can obtain and update the value mapped value at this physical page address through our virtual memory address.

\todo[inline]{Cite Kolanski here; not all of them, but some actually defined it exactly this way, for the current
address space.}

\begin{figure*}
  \[
  \begin{array}{l}
\begin{array}{l}
  \vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\vpage\; : \mathsf{vProp} \; \Sigma \;  \stackrel{{\triangle}}{:=} 
  \exists_{\entryf \;, \entrytr \;, \entrytw \;,\entryo} \ldotp

  \ulcorner \textsf{aligned } \vaddr \urcorner \ast  \\
 \quad \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr 
\end{array} \\
\\
\textsf{where   }\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \stackrel{\triangle}{=} \\
 \qquad\qquad \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qfracfotsss\naddr \ast \\ 
 \qquad\qquad  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qfracfotss\naddr  \ast  \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qfracfots\naddr \ast \\
  \qquad\qquad \nfpointsto{\mask\vaddr\ft\entrytw}{\mask\vaddr\tw\entrytw}\entryo\qfracfot\naddr 
   \end{array}
\]
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't that fraction be a q, not a 1, in the second line?}
  \label{fig:strongvirtualpointsto}
\end{figure*}
\end{remark}
%then we reach to a physical page addr (e.g. \textsf{pa}) 
%\[  \paddr\mapsto_{\textsf{a}}\{1\}\vsome \ast \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfrac\naddr \]

  \subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
  is too strong to specify some operations updating a page table entry.
   To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
  ensure fractional ownership of the four physical page table entries accessed by the hardware during address translation,
  as well as the physical memory holding the data of the virtual points-to.
  However, this formulation stores knowledge of the page table walk's details with ownership of the backing physical
  memory.
  Updating any of these mappings (e.g., move the page tables in physical memory, as in coallescing for
  superpages or hugepages)
  would require explicitly collecting all virtual points-to facts that traverse affected entries.
  It is preferable to permit the page tables themselves to be updated independently of the virtual points-to assertions,
  so long as those updates preserve the same virtual-to-physical translations.

\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{:=} 
    \exists \entryo \ldotp
    \sumapaces\rtv\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}  
  One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation. To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$. We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
 where
  \begin{itemize}
  \item we decorate the fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$) with the physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
  \item we place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership grants sound update on the page table entries
  \end{itemize}

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{:=} 
  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo)}
\end{array}
\]
\caption{Global Address-Space Invariant}
  \label{fig:peraspaceinvariant}
  \end{figure*}

@@

\[
  \begin{array}{l}
    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \mathsf{P},\; \mathsf{P r}. 
  \end{array}
  \]
  \label{fig:modaldef}
  \end{figure}
